### 2024-01-23 outputs/paper/paper.qmd
edited tables in the paper

---
title: "How Toronto's homeless death toll over the years shows the dangers of addiction's epidemic?"
subtitle: "Data analysis of homeless deaths in Toronto from 2017 to 2023"
author: 
  - Shipeng Zhang
thanks: "Code and data are available at: LINK."
date: 2024-01-22
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false
library("janitor")
library("knitr")
library("lubridate")
library("opendatatoronto")
library("tidyverse")
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| warning: false
#| message: false
#| echo: false

file_path <- "../../inputs/data/raw_data.csv"
toronto_homeless_death <- read.csv(file_path)
toronto_homeless_death <- dplyr::select(toronto_homeless_death, -X_id)           
sample_data <- head(toronto_homeless_death, 6)
knitr::kable(sample_data,'markdown',col.names = c("Year of Death", "Cause of Death", "Age Group", "Gender", "Count"),align = 'c',

                      caption = "Sample Table for Causes of Death")
  # Print the table

```

```{r}
#| label: fig-causes-of-death
#| fig-cap: Causes of death analysis
#| echo: false

ggplot(data = toronto_homeless_death,mapping =  aes(x = Cause_of_death, fill = Age_group)) +
  geom_bar(position = "stack") +
  scale_fill_manual(values = c("lightblue", "lightgreen", "lightpink", "lightyellow", "lightgray")) +
  labs(title = "Causes of Death Stacked by Age Group",
       x = "Cause of Death",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but
don't worry about doing that until you have finished every other aspect
of the paper - Quarto will try to make it look nice and the defaults
usually work well once you have enough text.)

#\| label: fig-planes #\| fig-cap: Relationship between wing length and
width #\| echo: false #\| warning: false #\| message: false

analysis_data \<- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data \|\> ggplot(aes(x = width, y = length)) + geom_point(alpha
= 0.8) + theme_minimal() + labs(x = "Wing width (mm)", y = "Wing length
(mm)")

```{r}

```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to
investigate... Background details and diagnostics are included in
[Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft.
Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both
measured in millimeters.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of
@rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time
spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance
$\theta$.

# Results

Our results are summarized in @tbl-modelresults. #\| echo: false #\|
eval: true #\| warning: false #\| message: false

library(rstanarm)

first_model \<- readRDS(file =
here::here("outputs/models/first_model.rds"))

```{r}

```

#\| echo: false #\| eval: true #\| label: tbl-modelresults #\| tbl-cap:
"Explanatory models of flight time based on wing width and wing length"
#\| warning: false

modelsummary::modelsummary( list( "First model" = first_model ),
statistic = "mad", fmt = 2 )

```{r}

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The
discussion is a chance to show off what you know and what you learnt
from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive
check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the
prior. This shows... #\| eval: true #\| echo: false #\| message: false
#\| warning: false #\| label: fig-ppcheckandposteriorvsprior #\|
layout-ncol: 2 #\| fig-cap: "Examining how the model fits, and is
affected by, the data" #\| fig-subcap: \["Posterior prediction check",
"Comparing the posterior with the prior"\]

pp_check(first_model) + theme_classic() + theme(legend.position =
"bottom")

posterior_vs_prior(first_model) + theme_minimal() +
scale_color_brewer(palette = "Set1") + theme(legend.position =
"bottom") + coord_flip()

```{r}

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests... #\|
echo: false #\| eval: true #\| message: false #\| warning: false #\|
label: fig-stanareyouokay #\| fig-cap: "Checking the convergence of the
MCMC algorithm" #\| fig-subcap: \["Trace plot", "Rhat plot"\] #\|
layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")

```{r}

```

\newpage

# References